SunPy - Python for Solar Physics

### 2014-01-24 3-1-VSO.tex

The Virtual Solar Observatory (VSO) was built to facilitate
access to solar data from the different archives around the 
world \cite{hill_virtual_2009}.  
Enabling therefore a single standard query interface to all
the available products from solar observation.  
VSO allows to query data not just by instrument or mission for 
a time range but by physical physical variables as for example:
intensity, magnetic field, particle flux or spectral range to name a few.
VSO's main interface is web-based\footnote{\url{http://virtualsolar.org}}, 
however a SOAP webservice is also available.
SunPy access to the VSO through the \texttt{vso} module which it uses 
the \texttt{suds} package to connect to the webservice.
This module was developed under the ESA's Summer of Code programme of 2010.

The \texttt{vso} module includes advances features through the use
of VSO attributes objects.  
The time range, the instrument, the wavelength, the physical observable,
and many other properties are set as attributes objects which could
be combine using \texttt{and} - just by chaining together the arguments -
or \texttt{or} operators - by joining the attributes using the ``$|$'' operator.
Listing~\ref{code:vso_query} shows few examples of queries made using
such attributes.

@@

The \texttt{vso} module also provides of an interface to download the data,
allowing to customise the directory structure to save the data.
Listing~\ref{code:vso_down} exemplifies such case.

>>>>

The Virtual Solar Observatory (VSO) was built to facilitate
access to solar data from the different archives around the 
world \cite{hill_virtual_2009}.  
Enabling therefore a single standard query interface to all
the available products from solar observation.  
VSO allows to query data not just by instrument or mission for 
a time range but by physical variables as for example:
intensity, magnetic field, particle flux or spectral range to name a few.
VSO's main interface is web-based\footnote{\url{http://virtualsolar.org}}, 
however a SOAP webservice is also available.
SunPy access to the VSO through the \texttt{vso} module which it uses 
the \texttt{suds} package to connect to the webservice.
This module was developed under the ESA's Summer of Code programme of 2010.

The \texttt{vso} module includes advanced features through the use
of VSO attributes objects.  
The time range, the instrument, the wavelength, the physical observable,
and many other properties are set as attributes objects which could
be combine using \texttt{and} - just by chaining together the arguments -
or \texttt{or} operators - by joining the attributes using the ``$|$'' operator.
Listing~\ref{code:vso_query} shows few examples of queries made using
such attributes.

@@

The \texttt{vso} module also provides an interface to download the data,
allowing to customise the directory structure to save the data.
Listing~\ref{code:vso_down} exemplifies such case.

### 2014-01-24 3-2-HEK.tex

SunPy accesses this information through the \texttt{hek} module, which
was developed through support from the European Space Agency Summer of
Code in Space (ESA-SOCIS) 2011.  The \texttt{hek} module makes use of
the HEK API
(http://vso.stanford.edu/hekwiki/ApplicationProgrammingInterface?action=print).
Every event type has required and optional properties; the current
list of events maintained by the HEK and their properties can be found
at http://www.lmsal.com/hek/VOEvent_Spec.html.

Simple HEK queries consist of start time, an end time, and an event
type (Example \ref{code:hek:simple}). Times can be specified in any
SunPy-acceptable format. Event types are specified as upper case, two
letter strings, and are identical to the two letter abbreviations
found at the HEK website, http://www.lmsal.com/hek/VOEvent_Spec.html.

\begin{listing}
\begin{minted}{python}
from sunpy.net import hek
client = hek.HEKClient()
tstart = '2011/08/09 07:23:56'
tend = '2011/08/09 12:40:29'
event_type = 'FL'
result = client.query(hek.attrs.Time(tstart,tend), hek.attrs.EventType(event_type))
\end{minted}
\caption{Simple HEK query, returning 19 entries.}
\label{code:hek:simple}
\end{listing}

@@

HEK attributes differ from VSO attributes (Section \ref{ssec:vso}) in
that many of them are wrappers that conveniently expose
comparisions by overloading Python operators.  This allows filtering
of the HEK entries by the properties of the event.  As was mentioned
above, the HEK stores feature/event metadata obtained in different
ways, known generally as {\it feature recognition methods}, or FRMs.
Example \ref{code:hek:frm} filters the results of the previous result
to return only those events that have the FRM 'SSW Latest Events'.
\begin{listing}
\begin{minted}{python}
result = client.query(hek.attrs.Time(tstart,tend), hek.attrs.EventType(event_type), hek.attrs.FRM.Name == 'SSW Latest Events')
\end{minted}
\caption{An HEK query that returns only those flares that were
  detected by the 'SSW Latest Events' feature recognition method.}
\label{code:hek:frm}
\end{listing}
Multiple comparisons can be made by including more comma-separated
conditions on the attributes in the call to the HEK query method.

HEK comparisons can be combined using Python’s logical operators AND
and OR. This makes complex queries easy to create: example
\ref{code:hek:or} returns flares west of 50 arcseconds or those that
have a peak flux above 1000.0 (the units of the flux are FRM dependent
and are described at the HEK website).
\begin{listing}
\begin{minted}{python}
result = client.query(hek.attrs.Time(tstart,tend), hek.attrs.EventType(event_type), (hek.attrs.Event.Coord1 > 50) or (hek.attrs.FL.PeakFlux > 1000.0) )
\caption{HEK query using the 'or' operator.}
\label{code:hek:or}
\end{listing}
All FRMs report the required feature attributes (as described at
http://www.lmsal.com/hek/VOEvent_Spec.html), but the optional
attributes are FRM dependent.  If a FRM does not have one of the
optional attribute, a 'None' is returned by the \texttt{hek} module.

>>>>

SunPy accesses this information through the \texttt{hek} module, which
was developed through support from the European Space Agency Summer of
Code in Space (ESA-SOCIS) 2011.  The \texttt{hek} module makes use of
the HEK API
(\url{http://vso.stanford.edu/hekwiki/ApplicationProgrammingInterface?action=print}).
Every event type has required and optional properties; the current
list of events maintained by the HEK and their properties can be found
at \url{http://www.lmsal.com/hek/VOEvent_Spec.html}.

Simple HEK queries consist of start time, an end time, and an event
type (Example \ref{code:hek:simple}). Times can be specified in any
SunPy-acceptable format. Event types are specified as upper case, two
letter strings, and are identical to the two letter abbreviations
found at the HEK website, \url{http://www.lmsal.com/hek/VOEvent_Spec.html}.

\begin{listing}[H]
\begin{minted}{python}
from sunpy.net import hek
client = hek.HEKClient()
tstart = '2011/08/09 07:23:56'
tend = '2011/08/09 12:40:29'
event\_type = 'FL'
result = client.query(hek.attrs.Time(tstart,tend), 
hek.attrs.EventType(event\_type))
\end{minted}
\caption{Simple HEK query, returning 19 entries.}
\label{code:hek:simple}
\end{listing}

@@

HEK attributes differ from VSO attributes (Section \ref{ssec:vso}) in
that many of them are wrappers that conveniently expose
comparisions by overloading Python operators.  This allows filtering
of the HEK entries by the properties of the event.  As was mentioned
above, the HEK stores feature/event metadata obtained in different
ways, known generally as {\it feature recognition methods}, or FRMs.
Example \ref{code:hek:frm} filters the results of the previous result
to return only those events that have the FRM 'SSW Latest Events'.
\begin{listing}[H]
\begin{minted}{python}
result = client.query(hek.attrs.Time(tstart,tend), 
hek.attrs.EventType(event\_type), hek.attrs.FRM.Name == 'SSW Latest Events')
\end{minted}
\caption{An HEK query that returns only those flares that were
  detected by the 'SSW Latest Events' feature recognition method.}
\label{code:hek:frm}
\end{listing}
Multiple comparisons can be made by including more comma-separated
conditions on the attributes in the call to the HEK query method.

HEK comparisons can be combined using Python’s logical operators AND
and OR. This makes complex queries easy to create: example
\ref{code:hek:or} returns flares west of 50 arcseconds or those that
have a peak flux above 1000.0 (the units of the flux are FRM dependent
and are described at the HEK website).
\begin{listing}[H]
\begin{minted}{python}
result = client.query(hek.attrs.Time(tstart,tend), 
hek.attrs.EventType(event\_type), (hek.attrs.Event.Coord1 > 50) or 
(hek.attrs.FL.PeakFlux > 1000.0) )
\end{minted}
\caption{HEK query using the 'or' operator.}
\label{code:hek:or}
\end{listing}
All FRMs report the required feature attributes (as described at
\url{http://www.lmsal.com/hek/VOEvent_Spec.html}), but the optional
attributes are FRM dependent.  If a FRM does not have one of the
optional attribute, a 'None' is returned by the \texttt{hek} module.

### 2014-01-24 3-3-HELIO.tex

The current version of SunPy has includes an interface to access the 
\textit{HELIO Event Catalogue} (HEC) service developed as part of Google
Summer of Code project in 2013.
This service provides at the moment access to 84 catalogues from different
sources (automated and manual classification, event lists from publications, \cdot).
HEC results (as the rest of HELIO services) provides the results in VOTable 
data format (defined by IVOA \cite{ochsenbein_ivoa_2011})
which the \textttt{astropy.io.votable} can interact with.
This format has the advantage of contain metadata with information like
data provenance and the performed query.

@@

Obtain from all the catalogues these which name contain our event of interest
>>> catalogues = hc.get_table_names()
>>> catalogues_event = [l[0] for l in catalogues if event_type in l[0]]
>>> print catalogues_event

Query all the catalogues that comes from cactus
>>> results = [hc.time_query(tstart, tend, event) for event in catalogues_event if 'cactus' in event]
>>> print results
********** This need to be added!
\end{minted}
\caption{Demonstration of the \texttt{helio.hec} interface.}
\label{code:helio}
\end{listing}

>>>>

The current version of SunPy has includes an interface to access the 
\textit{HELIO Event Catalogue} (HEC) service developed as part of Google
Summer of Code project in 2013.
This service provides at the moment access to 84 catalogues from different
sources (automated and manual classification, event lists from publications, \ldots).
HEC results (as the rest of HELIO services) provides the results in VOTable 
data format (defined by IVOA \cite{ochsenbein_ivoa_2011})
which the \texttt{astropy.io.votable} can interact with.
This format has the advantage of contain metadata with information like
data provenance and the performed query.

@@

Obtain from all the catalogues these which name contain our event of interest
>>> catalogues = hc.get_table_names()
>>> catalogues_event = [l[0] for l in catalogues if event_type in l[0]]
>>> print catalogues_event
******** This need to be added!

Query all the catalogues that comes from cactus
>>> results = [hc.time_query(tstart, tend, event) 
...            for event in catalogues_event if 'cactus' in event]
>>> print results
********** This need to be added!
\end{minted}
\caption{Demonstration of the \texttt{helio.hec} interface.}
\label{code:helio}
\end{listing}

### 2014-01-24 6-Future.tex

In concert with the work on the data types further integration with the 
\texttt{astropy} package will enable SunPy to inherit the large amounts of work 
done in that project to support a platform for astronomical data analysis. For 
example the \texttt{astropy.units} submodule provides a simple and efficient 
way of representing physical units in Python data types, there are plans to 
incorporate this widely in the SunPy code-base. In the past 12 months the 
collaboration with the Astropy project \cite{theastropycollaboration2013} and 
it's community has grown, this direct collaboration between Astropy and 
SunPy will be one of the key strengths of SunPy going forward.

>>>>

In concert with the work on the data types further integration with 
the 
\texttt{astropy} package will enable SunPy to inherit the large 
amounts of work 
done in that project to support a platform for astronomical data 
analysis. For 
example the \texttt{astropy.units} submodule provides a simple and 
efficient 
way of representing physical units in Python data types, there are 
plans to 
incorporate this widely in the SunPy code-base. In the past 12 months 
the 
collaboration with the Astropy project 
\citep{theastropycollaboration2013} and 
it's community has grown, this direct collaboration between Astropy 
and 
SunPy will be one of the key strengths of SunPy going forward.
